---
title: "Project Script"
output: html_notebook
---

This script loads in, cleans, and analyzes our project data from the course. 

Roadmap: 

0. Load in packages, metadata, and batch-processed file
0.5 Replace known names in batch-processed file
1. Clean batch-processed file according to John's tutorial
2. Qualitatively compare to manual file for aster and lupine
3. Visually check output file with built-in graphics
4. Run statistics on final output file
5. Create visuals for presentation

# 0. Load everything in

Install and load in *bouquet* package

```{r}
#install.packages("remotes")
remotes::install_github("jmpowers/bouquet", build_vignettes = TRUE)
```

```{r}
library(tidyverse)
library(bouquet)
library(ggplot2)
library(vegan)
```

```{r}
metadata <- read.csv("~/Desktop/GitHub/RMBLVolatile/Data/metadata.csv", header = TRUE)
metadata <- metadata[, c(1, 3:7)]

metadata <- load_metadata(metadata, sample = "file_name", date = "ambient.date", group = c("species", "replicate"), type = "sample_type", amount = "mass")

```

```{r}
source("read_shimadzu.R")
output <- read.shimadzu("~/Desktop/GitHub/RMBLVolatile/Data/Batchrun2.txt")

longdata <- load_longdata(output, sample = "Filename", RT = "Ret.Time", 
                          name = "Name", area = "Area", 
                          match = "SI", maxmatch=100)

longdataclean <- longdata %>%
  select(sample, RT, name, area, match)

```

# 0.5. DON'T USE: Replace known names and emission rates in longdataclean

```{r}
manualAst <- read.csv("~/Desktop/GitHub/RMBLVolatile/Data/manualAst.csv", header = TRUE)
manualLup <- read.csv("~/Desktop/GitHub/RMBLVolatile/Data/manualLup.csv", header = TRUE)
```

```{r}
manualAst <- manualAst %>%
  select(Ret.Time = Ret.Time, New.Name = Name, Em.Rate = X.1)

manualLup <- manualLup %>%
  select(Ret.Time = Ret.Time, New.Name = Name, Em.Rate = Adjust.for.weight..ng.hr.)
```

```{r}
longdataclean <- merge(longdataclean, manualAst, by.x = "RT", by.y = "Ret.Time", all.x = TRUE)
longdataclean <- merge(longdataclean, manualLup, by.x = "RT", by.y = "Ret.Time", all.x = TRUE)

longdataclean<- longdataclean %>%
  unite("New.Name", c("New.Name.x", "New.Name.y"), remove = TRUE , na.rm = TRUE)

longdataclean<- longdataclean %>%
  unite("Em.Rate", c("Em.Rate.x", "Em.Rate.y"), remove = TRUE, na.rm = TRUE)

```

## Get scaled emission rates for every peak in longdataclean

```{r}
longdataclean$Em.Rate.New <- 0

for(i in 1:length(longdataclean$RT)){
  if(longdataclean[i, 2] == "Aster1_floral_ALP.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 0.686
  }
  else if(longdataclean[i, 2] == "Aster2_floral_AFM.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 0.686 
  } 
  else if(longdataclean[i, 2] == "Aster3_floral_AFM.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 0.690
  }
  else if(longdataclean[i, 2] == "Lupine1_floral_AFM.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 1.025
  }
  else if(longdataclean[i, 2] == "Lupine2_floral_ALP.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 0.558
  } 
  else if(longdataclean[i, 2] == "Lupine3_floral_ALP.qgd"){
    longdataclean[i, 8] <- (30/(8274196/ longdataclean[i, 4])) / 0.919
  }
}
```


# 1. Clean batch file! 

## Make sampletable

```{r}
sampletable <- make_sampletable(longdataclean, metadata)
```


## Make chemtable

```{r}
chemtable <- make_chemtable(longdataclean, metadata)
```

## Filter out round 1

```{r}
#contaminants <- c("Tetratriacontane", "Tetratetracontane", "Tetradecanoic acid", "Tetracosane", "Silanediol, dimethyl-", "Phthalic acid, 7-bromoheptyl butyl ester", "Phosphonic acid, (p-hydroxyphenyl)-", "Pentadecanal-", "Pentacontanoic acid, ethyl ester", "Oxybenzone", "n-Tridecan-1-ol", "n-Tetracosanol-1", "n-Pentadecanol", "n-Nonadecanol-1", "n-Heptadecanol-1", "Methyl stearate", "Hexanoic acid")

chemtable <- chemtable %>%
  filter_freq(0.66, group = TRUE) %>%
  filter_RT(4, 18) %>%
  filter_area(min_maximum = 5e5) %>%
  filter_ambient_ratio(sampletable, metadata, ratio = 3) %>%
 #filter_contaminant(cont.list = contaminants) %>%
  filter_ambient_date(sampletable, metadata, ratio = 3, prop_dates = .1)
```

## Make preliminary final filter

```{r}
chemtable <- chemtable %>% 
  combine_filters() %>% 
  within(filter_final <- 
              ((filter_freq.Heliomeris_multiflora == "OK" | filter_freq.Lupinus_argenteus == "OK")
               & filter_area == "OK" & filter_ambient_date == "OK" & filter_RT == "OK" & filter_ambient_ratio == "OK"))

```

## Diagnose what filters are doing what

```{r}
plot_filters(chemtable, option = "prop")
```

```{r}
plot_filters(chemtable, option = "rarity", yrange=2.5, fontsize=2.5, pointsize=15)
```


```{r}
plot_filters(chemtable, option = "ambient", yrange=4, pointsize=6)
```

## Subset final dataset by filter values 

```{r}
finaltable <- prune_sampletable(sampletable, chemtable, metadata)
finalmetadata <- subset(metadata, type == "floral")

write.csv(file = "~/Desktop/GitHub/RMBLVolatile/Data/finaltable.csv", finaltable)
```

# 4. Run statistics

```{r}
permanova <- adonis2(finaltable ~ species, data = finalmetadata)

permanova

cap <- capscale(finaltable ~ species, data = finalmetadata, 
                 distance = "bray")
```

# 5. Graphing


```{r}
plot(cap, type = "n", 
     main = paste("CAP of", ncol(finaltable), "volatiles in", 
                  nrow(finaltable), "samples vs. species"))
legend("bottomright", levels(finalmetadata$species), 
       col = 2:5, pch = 19, title = "Species")
points(cap, col = as.integer(finalmetadata$species)+1, pch = 19)
text(cap, display = "species", col="grey50")
text(cap, display = "cn", labels=levels(finalmetadata$species))

```


```{r}
heatmap(as.matrix(finaltable)^(1/4), 
        scale = "none", distfun = vegdist,
        cexCol = 1, margins = c(15, 0), labRow = NA,
        col=colorRampPalette(c("white","darkblue"))(256),
        RowSideColors = as.character(as.integer(finalmetadata$species)+1))
legend("bottomleft", levels(finalmetadata$species), title = "Species",
       col = 2:5, pch = 19, inset=c(0,-0.2),  xpd=T)
```


# SECTION 2: Rerun with just the ones that we IDed from our individuals

```{r}
longdatapruned <- longdataclean[longdataclean$New.Name != "", ]
```

**Retention times were not close enough to do this thing between runs** 

Instead, we are going to combine the manual datafiles and make a pretty graph from those. 

## Make a new datafile with the correct column names

```{r}
manualAst2 <- read.csv("~/Desktop/GitHub/RMBLVolatile/Data/manualAst.csv", header = TRUE)
manualLup2 <- read.csv("~/Desktop/GitHub/RMBLVolatile/Data/manualLup.csv", header = TRUE)

manualAst2$Sample <- "Aster1_floral_ALP.qgd"
manualLup2$Sample <- "Lupine3_floral_ALP.qgd"
manualAst2$match <- 1
manualLup2$match <- 1

manualAst2 <- manualAst2 %>%
  select(RT = Ret.Time, sample = Sample, name = Name, area = X.1, match = match)

manualLup2 <- manualLup2 %>%
  select(RT = Ret.Time, sample = Sample, name = Name, area = Adjust.for.weight..ng.hr., match = match)

newlongdata <- rbind(manualAst2, manualLup2)
```




